The evolution of gene expression in the term placenta of viviparous mammals

### 2016-03-10 manuscript/manuscript.Rnw
add changes Michael requested

@@

The placenta arose at some point before the origin of extant mammals,
a group that includes marsupials, monotremes, and eutherian or
“placental mammals”. The term “placental mammals” is misleading, as
extant yinotherians (such as \textit{Ornithorhynchus anatinus}, the
platypus) possess a rudimentary placenta, as do marsupials, although
the direct apposition to maternal tissues lasts for a proportionally
shorter time during gestation
\cite{Selwood.ea2006:Trophoblastandhypoblastin,Renfree2010:ReviewMarsupialsplacentalmammals}.
Marsupials also differ from eutherian mammals in that most possess
only a choriovitelline placenta, produced from the fusion of the
chorion and yolk sac, during most of gestation. Although some
eutherian mammals may possess a choriovitelline placenta at some point
in development, the major site of nutrient exchange in placentals
occurs via the chorioallantoic placenta, produced from the fusion of
the chorion and allantois
\cite{Mossman.1987:vertebratefetalmembranes}. It is unclear whether
the first placenta was more like that of marsupials or eutherians;
however, using phylogenetic methods, the ancestral eutherian was
reconstructed as possessing a relatively invasive hemochorial
placenta, in which the maternal and fetal blood supply is separated by
one tissue layer, as in humans and mouse
\cite{Elliot.ea2009:Phylogeneticevidenceforearly,Wildman.ea2006:Evolutionofmammalianplacenta}.
\DLA{To discuss maternal/fetal split with derek}

Although the morphology of the mammalian placenta has been well
characterized in a number of species
\cite{Mossman.1987:vertebratefetalmembranes}, the diversity of the
molecular environment is only beginning to be understood from a
genetic perspective \cite{Carter2012:Evolutionofplacentalfunction}.
Studies of certain model species (i.e., \textit{Homo sapiens},
\textit{Mus musculus}, \textit{Bos taurus}) have revealed a stunning
divergence in the genetics of the placenta, especially in later stages
of pregnancy
\cite{Carter2012:Evolutionofplacentalfunction,Hou.ea2012:Elephanttranscriptomeprovidesinsights,Knox.ea2008:Genomicevolutionofplacenta,Su.ea2002:Large-scaleanalysisofhuman}.
For example, each lineage has evolved expansive placenta-specific gene
families with expression patterns that can vary in space and time
\cite{Green.ea2000:Pregnancy-associatedbovineandovine,Simmons.ea2008:Spatialandtemporalexpression}.
These include independent expansions of the prolactins in \mouse,
\textit{Spalax carmeli}, \textit{Spalax galili}, and \cow, \acp{PSG} in \human and \mouse, \acp{PAG}
in \cow, genes related to the thymus-specific serine protease
\textit{PRSS16} in \textit{Loxodona africana}, and the expansion of the growth hormone
and chorionic gonadotropin beta genes in hominoid primates
\cite{Haig2008:Placentalgrowthhormone-relatedproteins,Rawn.ea2008:evolutionregulationandfunction}.
Expression patterns of specific genes also can vary widely between
species. For example, the expression of orthologous genes in the
placenta have been found to vary substantially between \human and
\mouse \cite{Su.ea2002:Large-scaleanalysisofhuman}, and the
recruitment of \textit{SIGLEC6} to placental expression in \human
differs from both \textit{Pan troglodytes} and \mouse
\cite{Linden.ea2007:Human-specificexpressionofSiglec-6}.

@@

One symptom of this maternal-fetal conflict is its potential affect on
health
\cite{Hollegaard.ea2013:Parent-offspringconflictandpersistence}. In
addition to having an invasive hemochorial placenta, \human possess a
deeply invasive trophoblast layer that can penetrate the myometrium of
the uterus \cite{Carter.ea2011:Evolutionofinvasiveplacentation}.
Defects in placental growth are known to be associated with the
devastating condition of pre-eclampsia, which can cause hypertension,
proteinuria, and has the potential to lead to death of the mother
and/or infant \cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}.
Although cases have been documented in \textit{P. troglodytes}
\cite{Stout.ea1969:Glomerularcapillaryendothelialswelling}, and
\textit{Gorilla gorilla}
\cite{Thornton.ea1992:Convulsionsinpregnancyin}, pre-eclampsia has a
far higher incidence in \human pregnancies with as much as 4\% of
pregnancies being affected in the United States
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}. \textit{P.
  troglodytes} and \textit{G. gorilla} also have deeply invasive
trophoblasts, although the progression of invasion in great apes may
be slower than \human
\cite{Carter.ea2011:Evolutionofinvasiveplacentation}; however,
large-scale expression studies of great ape placentas have not been
conducted. The increasing affordability of next-generation sequencing
technology has enabled the large-scale study of specific tissues from
a wide array of species to address broader phylogenetic questions in
the evolution of gene expression
\cite{Brawand.ea2011:evolutionofgeneexpression}. Here we use RNA-Seq
technology to sequence and quantify the mRNA environment of mammalian
placentas at term, including that of the \human, \textit{Pan paniscus}
(the sister species to \textit{P. troglodytes}), \textit{Ateles
  fusciceps}, \mouse, \galili, \carmeli, \textit{Canis familiaris},
\cow, \elephant, \textit{Dasypus novemcinctus}, and
\textit{Monodelphis domesticus}. These species represent four major
branches of eutherian mammals (Euarchontoglires, Laurasiatheria,
Afrotheria, and Xenartha) as well as marsupials. These placentas also represent a
wide array of placental morphologies as well as traits such as number
and size of offspring and length of gestation
(\cref{tab:gestational_periods}). We use this data, along with
existing RNA-Seq data from \mouse
\cite{Wang.ea2011:surveyfornovelimprinted}, \human
\cite{Hughes.ea2015:Evaluatingintra-andinter-individual}, \textit{Equus caballus}
\cite{Coleman.ea2010:Structuralannotationofequine}, \textit{Ovis aries}
\cite{Archibald.ea2010:sheepgenomereferencesequence}, and \textit{Sus scrofa}
\cite{Ahanda.ea2012:Predictionofaltered3-} to address the major
differences between expression in marsupial and eutherian mammals at
term, identify the evolution of expression in important placental
genes such as hormones and galectins, identify novel lineage-specific
expression patterns, and quantify \human specific differences in gene
expression that may be related to obstetrical disorders such as
pre-eclampsia.

@@

\subsection{Correlation of Tissue Specific Index with Lineage Specific Genes}

To evaluate the extent of tissue specificity of each gene, we
calculated the \ac{TSI} based on human RNASeq data in 3267 samples in
126 tissues the Roadmap Epigenomics
\cite{Bernstein.ea2010:NIHRoadmapEpigenomicsMapping} and GTEx projects
\cite{Gtexconsortium.2015:Thegenotypetissueexpression} as well
additional placenta
\cite{Hughes.ea2015:Evaluatingintra-andinter-individual} and uterus
\cite{Chan.ea2014:Assessmentofmyometrialtranscriptome} samples from
SRA. The \ac{TSI} measures the extent to which a specific gene has
predominant expression in only one tissue
\cite{Yanai.ea2005:Genome-widemidrangetranscriptionprofiles}. A gene
expressed in only one tissue would have a \ac{TSI} of 1, whereas a
gene expressed in all tissues would have a \ac{TSI} of 0.

@@

<<load.placenta.lineage.specific,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("placenta_lineage_specific")))>>=
load("../data/placenta_lineage_specific")
source("/home/don/projects/xtable/pkg/R/print.xtable.R")

placenta.lineage.specific.table <- function(analysis,output=TRUE) {
    temp.table <- 
        head(placenta.lineage.specific.analysis[analysis.name==analysis &
                                                p.value <= 0.05][order(p.value),
                                                                 list(human_name,p.value,fdr,fc),],
             n=5)
    temp.table <- data.table(temp.table)
    setkey(temp.table,"human_name")
    temp.table <- 
        hugo.nomenclature[temp.table][,list(`Approved Symbol`,
                                            `Approved Name`,p.value,fdr,fc)]
    setnames(temp.table,
              c("Human Symbol","Human Name","p (t-test)","FDR","Fold Change")
              )
    setkey(temp.table,"Human Symbol")
    if (output) {
        display.table <- temp.table[order(`p (t-test)`),]
        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        print(xtable(display.table,
                     display=c("s","s","s","e","g","f"),
                     ),
              booktabs=TRUE,
              floating=FALSE,
              math.style.exponents=TRUE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)

@@

@@

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of \human are shown in
\cref{tab:top50_human_placenta} with all genes with \ac{FPKM} $\ge 50$
in Supplemental \cref{tab:all_homo_sapiens_placenta}. The gene with by far
the highest \ac{FPKM} value
%
(\Sexpr{gene.expression(species.name="homo sapiens",gene="CSH1")}) %
%
in \human is \textit{CSH1}, followed by \textit{CSH2}, and \textit{CSHL1}, all
three of which are members of a distinct expansion of the growth
hormone gene family within hominoids (\cref{fig:galectins_evolution_tree}). Another member of the growth
hormone gene family, \ac{GH2} (\ac{FPKM}=%
\Sexpr{gene.expression(species.name="homo sapiens",gene="GH2")}), %
is also expressed to a high degree in the \human placenta
(\cref{tab:top50_human_placenta}).
% 
Eleven \acp{PSG} are highly-expressed
(\ac{FPKM}$\ge 100$) in the term placenta of \human, with \textit{PSG1} showing the
greatest degree of expression %
(\Sexpr{gene.expression(species.name="homo sapiens",gene="PSG1")}) % 
and the pseudogene \textit{PSG10P} showing the least %
(\Sexpr{gene.expression(species.name="homo sapiens",gene="PSG10P")}, \cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~A) . %
These
genes are members of a hominoid-specific expansion of the \ac{CEA} family
and known to be expressed by the syncytiotrophoblast
\cite{Rawn.ea2008:evolutionregulationandfunction,Lei.ea1992:Cloningandexpressionof}. %
%
Genes involved in the production of estrogens (%
\Sexpr{gene.expression(species.name="homo sapiens",gene="CYP19A1",complete=TRUE)},
% 
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD17B1",complete=TRUE)}, and
% 
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD3B1",complete=TRUE)}%
% 
) and progesterone (%
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD3B1",complete=TRUE)},
\Sexpr{gene.expression(species.name="homo sapiens",gene="CYP11A1",complete=TRUE)}, and
\Sexpr{gene.expression(species.name="homo sapiens",gene="ADM",complete=TRUE)})
%
show exceptionally high expression in the human placenta, as well as %
%
\textit{CGA} (FPKM=\Sexpr{gene.expression(species.name="homo sapiens",gene="CGA")}, %
%
which generates the alpha subnit of multiple human chorionic
gonadotropin hormones (\cref{tab:top50_human_placenta}). The genes
producing the beta subunits of the chorionic gonadotropins
(\textit{CGB}, \textit{CGB1}, \textit{CGB2}, \textit{CGB5},
\textit{CGB7}, \textit{CGB8}) have reasonable expression
($\mathrm{FPKM} \ge \Sexpr{cgb.min.expression}$), and the sum of their
expression is on the order of the expression of expression of
\textit{CGA} ($\sum \mathit{CGBs}=\Sexpr{cgb.sum.expression}$ FPKM) .
%
%
Other highly expressed genes in the top 50 include members of the
hemoglobin family (e.g. \textit{HBA1}, \textit{HBB}, \textit{HBG2},
\textit{HBA2}), the maternally imprinted long noncoding RNA gene
\textit{H19}, genes coding for proteins required for growth and/or
cell proliferation (\textit{ADAM12}, \textit{KISS1}, \textit{IGF2},
\textit{GPC3}, \textit{CSF3R}), anticoagulation (\textit{TFPI2}), and
glucose transport (\textit{SLC2A1}) among others
(\cref{tab:top50_human_placenta}, Supplemental
\cref{tab:all_homo_sapiens_placenta}).

@@

We examined the expression of three rodentia, \mouse, \carmeli, and
\galili{}. We lacked sufficient power to detect any significant
changes on the Rodentia lineage in comparison to all other analyzed
species (Supplemental \cref{tab:rodentia_lineage_specific}).

@@

@@

>>>>

  \DLA{Complete rewrite abstract to make the points that we have found
    stronger and less weak.}

@@

The placenta arose at some point before the origin of extant mammals,
a group that includes marsupials, monotremes, and eutherian or
“placental mammals”. The term “placental mammals” is misleading, as
extant yinotherians (such as \textit{Ornithorhynchus anatinus}, the
platypus) possess a rudimentary placenta, as do marsupials, although
the direct apposition to maternal tissues lasts for a proportionally
shorter time during gestation
\cite{Selwood.ea2006:Trophoblastandhypoblastin,Renfree2010:ReviewMarsupialsplacentalmammals}.
Marsupials also differ from eutherian mammals in that most possess
only a choriovitelline placenta, produced from the fusion of the
chorion and yolk sac, during most of gestation. Although some
eutherian mammals may possess a choriovitelline placenta at some point
in development, the major site of nutrient exchange in placentals
occurs via the chorioallantoic placenta, produced from the fusion of
the chorion and allantois
\cite{Mossman.1987:vertebratefetalmembranes}. It is unclear whether
the first placenta was more like that of marsupials or eutherians;
however, using phylogenetic methods, the ancestral eutherian was
reconstructed as possessing a relatively invasive hemochorial
placenta, as in humans and mouse
\cite{Elliot.ea2009:Phylogeneticevidenceforearly,Wildman.ea2006:Evolutionofmammalianplacenta}.

Although the morphology of the mammalian placenta has been well
characterized in a number of species
\cite{Mossman.1987:vertebratefetalmembranes}, the diversity of the
molecular environment is only beginning to be understood from a
genetic perspective \cite{Carter2012:Evolutionofplacentalfunction}.
Studies of certain model species (i.e., \textit{Homo sapiens},
\textit{Mus musculus}, \textit{Bos taurus}) have revealed a stunning
divergence in the genetics of the placenta, especially in later stages
of pregnancy
\cite{Carter2012:Evolutionofplacentalfunction,Hou.ea2012:Elephanttranscriptomeprovidesinsights,Knox.ea2008:Genomicevolutionofplacenta,Su.ea2002:Large-scaleanalysisofhuman}.
For example, each lineage has evolved expansive placenta-specific gene
families with expression patterns that can vary in space and time
\cite{Green.ea2000:Pregnancy-associatedbovineandovine,Simmons.ea2008:Spatialandtemporalexpression}.
These include independent expansions of the prolactins in \mouse,
\textit{Spalax carmeli}, \textit{Spalax galili}, and \cow, \acp{PSG} in \human and \mouse, \acp{PAG}
in \cow, genes related to the thymus-specific serine protease
\textit{PRSS16} in \textit{Loxodona africana} (in this work),
 and the expansion of the growth hormone
and chorionic gonadotropin beta genes in hominoid primates
\cite{Haig2008:Placentalgrowthhormone-relatedproteins,Rawn.ea2008:evolutionregulationandfunction}.
Expression patterns of specific genes also can vary widely between
species. For example, the expression of orthologous genes in the
placenta have been found to vary substantially between \human and
\mouse \cite{Su.ea2002:Large-scaleanalysisofhuman}, and the
recruitment of \textit{SIGLEC6} to placental expression in \human
differs from both \textit{Pan troglodytes} and \mouse
\cite{Linden.ea2007:Human-specificexpressionofSiglec-6}.

@@

One symptom of this maternal-fetal conflict is its potential affect on
health
\cite{Hollegaard.ea2013:Parent-offspringconflictandpersistence}. In
addition to having an invasive hemochorial placenta, \human possess a
deeply invasive trophoblast layer that can penetrate the myometrium of
the uterus \cite{Carter.ea2011:Evolutionofinvasiveplacentation}.
Defects in placental growth are known to be associated with the
devastating condition of pre-eclampsia, which can cause hypertension,
proteinuria, and has the potential to lead to death of the mother
and/or infant \cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}.
Although cases have been documented in \textit{P. troglodytes}
\cite{Stout.ea1969:Glomerularcapillaryendothelialswelling}, and
\textit{Gorilla gorilla}
\cite{Thornton.ea1992:Convulsionsinpregnancyin}, pre-eclampsia has a
far higher incidence in \human pregnancies with as much as 4\% of
pregnancies being affected in the United States
\cite{Chaiworapongsa.ea2014:Pre-eclampsiapart1current}. \textit{P.
  troglodytes} and \textit{G. gorilla} also have deeply invasive
trophoblasts, although the progression of invasion in great apes may
be slower than \human
\cite{Carter.ea2011:Evolutionofinvasiveplacentation}. However,
large-scale expression studies of great ape placentas have not been
conducted. The increasing affordability of next-generation sequencing
technology has enabled the large-scale study of specific tissues from
a wide array of species to address broader phylogenetic questions in
the evolution of gene expression
\cite{Brawand.ea2011:evolutionofgeneexpression}.

Here we use RNA-Seq
technology to sequence and quantify the mRNA environment of mammalian
placentas at term, including that of the \human, \textit{Pan paniscus}
(the sister species to \textit{P. troglodytes}), \textit{Ateles
  fusciceps}, \mouse, \galili, \carmeli, \textit{Canis familiaris},
\cow, \elephant, \textit{Dasypus novemcinctus}, and
\textit{Monodelphis domesticus}. These species represent four major
branches of eutherian mammals (Euarchontoglires, Laurasiatheria,
Afrotheria, and Xenartha) as well as marsupials. These placentas also represent a
wide array of placental morphologies as well as traits such as number
and size of offspring and length of gestation
(\cref{tab:gestational_periods}). We use this data, along with
existing RNA-Seq data from \mouse
\cite{Wang.ea2011:surveyfornovelimprinted}, \human
\cite{Hughes.ea2015:Evaluatingintra-andinter-individual}, \textit{Equus caballus}
\cite{Coleman.ea2010:Structuralannotationofequine}, \textit{Ovis aries}
\cite{Archibald.ea2010:sheepgenomereferencesequence}, and \textit{Sus scrofa}
\cite{Ahanda.ea2012:Predictionofaltered3-} to address the major
differences between expression in marsupial and eutherian mammals at
term, identify the evolution of expression in important placental
genes such as hormones and galectins, identify novel lineage-specific
expression patterns, and quantify \human specific differences in gene
expression that may be related to obstetrical disorders such as
pre-eclampsia.

@@

% \subsection{Correlation of Tissue Specific Index with Lineage Specific Genes}
% 
% To evaluate the extent of tissue specificity of each gene, we
% calculated the \ac{TSI} based on human RNASeq data in 3267 samples in
% 126 tissues the Roadmap Epigenomics
% \cite{Bernstein.ea2010:NIHRoadmapEpigenomicsMapping} and GTEx projects
% \cite{Gtexconsortium.2015:Thegenotypetissueexpression} as well
% additional placenta
% \cite{Hughes.ea2015:Evaluatingintra-andinter-individual} and uterus
% \cite{Chan.ea2014:Assessmentofmyometrialtranscriptome} samples from
% SRA. The \ac{TSI} measures the extent to which a specific gene has
% predominant expression in only one tissue
% \cite{Yanai.ea2005:Genome-widemidrangetranscriptionprofiles}. A gene
% expressed in only one tissue would have a \ac{TSI} of 1, whereas a
% gene expressed in all tissues would have a \ac{TSI} of 0.

@@

<<load.placenta.lineage.specific,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("placenta_lineage_specific","human_mouse_lineage_specific")))>>=
load("../data/placenta_lineage_specific")
load("../data/human_mouse_lineage_specific")
source("/home/don/projects/xtable/pkg/R/print.xtable.R")

placenta.lineage.specific.table <- function(analysis,output=TRUE,lineage.table=placenta.lineage.specific.analysis,members=5) {
    temp.table <- 
        head(lineage.table[analysis.name==analysis &
                           p.value <= 0.05][order(p.value),
                                            list(human_name,p.value,fdr,fc),],
             n=members)
    temp.table <- data.table(temp.table)
    setkey(temp.table,"human_name")
    temp.table <- 
        hugo.nomenclature[temp.table][,list(`Approved Symbol`,
                                            `Approved Name`,p.value,fdr,fc)]
    setnames(temp.table,
              c("Human Symbol","Human Name","p (t-test)","FDR","Fold Change")
              )
    setkey(temp.table,"Human Symbol")
    if (output) {
        display.table <- temp.table[order(`p (t-test)`),]
        display.table <-
            display.table[,`Human Symbol`:=paste0("\\textit{",xtable.sanitize.text(`Human Symbol`),
                                                  "}")]
        print(xtable(display.table,
                     display=c("s","s","s","e","g","f"),
                     align=c("c","c","p{2in}","r","r","r")
                     ),
              booktabs=TRUE,
              floating=FALSE,
              math.style.exponents=TRUE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)

@@

\begin{table}
  \footnotesize
<<human.lineage.specific,echo=FALSE,results="asis">>=
hum.all.lin.spec <-
    placenta.lineage.specific.table("hum.all",
                                    lineage.table=human.mouse.linage.specific.analysis,
                                    members=40)
setkey(hum.all.lin.spec,"Human Symbol")
@ 
\caption{Genes whose expression changed significantly in \human
  as compared to all other analyzed species.}
\label{tab:human_lineage_specific}
\end{table}

@@

The top 50 expressed genes as measured by \ac{FPKM} in the term
placenta of \human are shown in
\cref{tab:top50_human_placenta} with all genes with \ac{FPKM} $\ge 50$
in Supplemental \cref{tab:all_homo_sapiens_placenta}. The gene with by far
the highest \ac{FPKM} value
%
(\Sexpr{gene.expression(species.name="homo sapiens",gene="CSH1")}) %
%
in \human is \textit{CSH1}, followed by \textit{CSH2}, and \textit{CSHL1}, all
three of which are members of a distinct expansion of the growth
hormone gene family within hominoids (\cref{fig:galectins_evolution_tree}). Another member of the growth
hormone gene family, \ac{GH2} (\ac{FPKM}=%
\Sexpr{gene.expression(species.name="homo sapiens",gene="GH2")}), %
is also expressed to a high degree in the \human placenta
(\cref{tab:top50_human_placenta}).
% 
Eleven \acp{PSG} are highly-expressed
(\ac{FPKM}$\ge 100$) in the term placenta of \human, with \textit{PSG1} showing the
greatest degree of expression %
(\Sexpr{gene.expression(species.name="homo sapiens",gene="PSG1")}) % 
and the pseudogene \textit{PSG10P} showing the least %
(\Sexpr{gene.expression(species.name="homo sapiens",gene="PSG10P")}, \cref{fig:pag2_prss16_ctsl_psg_evolution_tree}~A) . %
These
genes are members of a hominoid-specific expansion of the \ac{CEA} family
and known to be expressed by the syncytiotrophoblast
\cite{Rawn.ea2008:evolutionregulationandfunction,Lei.ea1992:Cloningandexpressionof}. %
%
Genes involved in the production of estrogens (%
\Sexpr{gene.expression(species.name="homo sapiens",gene="CYP19A1",complete=TRUE)},
% 
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD17B1",complete=TRUE)}, and
% 
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD3B1",complete=TRUE)}%
% 
) and progesterone (%
\Sexpr{gene.expression(species.name="homo sapiens",gene="HSD3B1",complete=TRUE)},
\Sexpr{gene.expression(species.name="homo sapiens",gene="CYP11A1",complete=TRUE)}, and
\Sexpr{gene.expression(species.name="homo sapiens",gene="ADM",complete=TRUE)})
%
show exceptionally high expression in the human placenta, as well as %
%
\textit{CGA} (FPKM=\Sexpr{gene.expression(species.name="homo sapiens",gene="CGA")}, %
%
which generates the alpha subnit of multiple human chorionic
gonadotropin hormones (\cref{tab:top50_human_placenta}). The genes
producing the beta subunits of the chorionic gonadotropins
(\textit{CGB}, \textit{CGB1}, \textit{CGB2}, \textit{CGB5},
\textit{CGB7}, \textit{CGB8}) have reasonable expression
($\mathrm{FPKM} \ge \Sexpr{cgb.min.expression}$), and the sum of their
expression is on the order of the expression of
\textit{CGA} ($\sum \mathit{CGBs}=\Sexpr{cgb.sum.expression}$ FPKM) .
%
%
Other highly expressed genes in the top 50 include members of the
hemoglobin family (e.g. \textit{HBA1}, \textit{HBB}, \textit{HBG2},
\textit{HBA2}), the maternally imprinted long noncoding RNA gene
\textit{H19}, genes coding for proteins required for growth and/or
cell proliferation (\textit{ADAM12}, \textit{KISS1}, \textit{IGF2},
\textit{GPC3}, \textit{CSF3R}), anticoagulation (\textit{TFPI2}), and
glucose transport (\textit{SLC2A1}) among others
(\cref{tab:top50_human_placenta}, Supplemental
\cref{tab:all_homo_sapiens_placenta}).

\DLA{Discuss pre-eclampsia specific genes here to make this point stronger}

@@

We examined the expression of three rodents, \mouse, \carmeli, and
\galili{}. We did not detect any significant changes on the Rodentia
lineage in comparison to all other analyzed species (Supplemental
\cref{tab:rodentia_lineage_specific}), likely because of a lack power
to overcome multiple testing.

@@

\begin{table}
<<mouse.lineage.specific,echo=FALSE,results="asis">>=
mou.all.lin.spec <-
    placenta.lineage.specific.table("mou.all",
                                    lineage.table=human.mouse.linage.specific.analysis)
setkey(mou.all.lin.spec,"Human Symbol")
@ 
\caption{Genes whose expression changed significantly in \mouse
  as compared to all other analyzed species.}
\label{tab:mouse_lineage_specific}
\end{table}

@@

\DLA{Make the discussion of pre-eclampsia clearer and more defined;
  possibly include some of the discuss below. Also, this is not
  necessarily a shift on the human line.}
